Logsums

### 2020-05-05 01_health_data.Rmd
Clean up file names and headers

---
title: "Health Data"
author: "Greg Macfarlane"
date: "6/6/2019"
output: html_document
---

@@

cdc500 <- lapply(cdc_urls, function(url){
  read.socrata(url, app_token = Sys.getenv("CDC_KEY"),
               stringsAsFactors = FALSE) %>%
    tbl_df() %>%
    filter(tractfips %in% acs$geoid) %>%
    transmute( geoid = tractfips, value = as.numeric(data_value ), year = year)
}) %>%
  bind_rows(.id = "measure") %>%
  spread(measure, value) %>%
  mutate(physact = 100 - physact)
```

>>>>

---
title: "Health Data"
---

@@

cdc500 <- lapply(cdc_urls, function(url){
  read.socrata(url, app_token = Sys.getenv("CDC_KEY"),
               stringsAsFactors = FALSE) %>%
    as_tibble() %>%
    filter(tractfips %in% acs$geoid) %>%
    transmute( geoid = tractfips, value = as.numeric(data_value ), year = year)
}) %>%
  bind_rows(.id = "measure") %>%
  spread(measure, value) %>%
  mutate(physact = 100 - physact)
```

### 2020-05-05 04_accessibility.Rmd
Clean up file names and headers

---
title: "Logsums"
author: "Greg Macfarlane"
date: "6/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(sf)
library(leaflet)
library(lubridate)
library(VGAM)
```

In this document we calculate the utility-based measure of accessibility, as well
as the 10-minute walk buffer.

## Distances
We cleaned and gathered the data parks data and data for tracts in the previous
two pages, as well as computed the network-based distance between each block
group and the nearest edge point of each park.

```{r load, message=FALSE}
open_spaces <- read_rds("data/open_spaces.rds")
blockgroups <- read_csv("data/blockgroups.csv")

# tigris not working, manual download
#bg_poly <- tigris::block_groups("New York", county =  c("081", "047", "061", "005", "085"))
bg_poly <- st_read("/Users/gregmacfarlane/Library/Caches/tigris/tl_2019_36_bg.shp") %>%
  filter(GEOID %in% blockgroups$GEOID) %>%
  transmute(geoid = GEOID) %>%
  st_transform(32618)
```

In calculating the distances, two things could happen. First, it is possible that
no path is found between the park and the block group centroid; this could be the
result of network miscodings or other errors in the network algorithm. We'll have
to take the results as they come, and we simply remove the park from that block
group's accessibility calculation. The other issue is when a block group centroid
and the park edge point match to the *same* node, resulting in a distance of zero.
Because we are going to log-transform the distance, we add ten meters to all paths.

```{r distance_adjust}
distance_files <- list.files("data/nyc_shortest_paths_3500m/", full.names = TRUE)

read_all_files <- function(pattern = pattern, path = "./"){
  map_df(list.files(path = path, pattern = pattern, full.names = TRUE), read_csv,
         col_types = list(geoid = col_character(), park_id = col_character()))
}
distances <- read_all_files(".csv", "data/nyc_shortest_paths_3500m/") %>%
  # filter out infinite paths
  filter(!is.infinite(distance)) %>%
  # add ten meters to all paths
  mutate(distance = distance + 10)
```

Now we can calculate the utility specification. The first specification uses the
distance and size terms from the Alameda County study. The second specification
adds the amenities with the utilities found by the Kinnell et al. study.

```{r logsums}
logsums <- distances %>%
  left_join(open_spaces %>% st_set_geometry(NULL), by = c("park_id" = "id")) %>%
  mutate(
    distance_miles = (distance + 10) * 0.000621371,
    u_access = -1.768 * log(distance_miles) + 0.373 * log(as.numeric(acres)),
    u_multi = u_access + 
      open_spaces$trails * 0.99 + 
      open_spaces$playgrounds * 0.26 + 
      open_spaces$courts * 0.43
  ) %>%
  group_by(geoid) %>%
  summarise(
    access_ls = log(sum(exp(u_access), na.rm = TRUE)),
    multi_ls = log(sum(exp(u_multi), na.rm = TRUE))
  ) %>%
  mutate_at(
    vars(ends_with("_ls")),
    list(~ (. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE))
  )

bg_access <- bg_poly %>%
  left_join(logsums, by = "geoid")
```

## 10-minute buffer

We are also interested in looking at what happens when the tracts are within a 
10-minute walk of a park, the threshold established by the Trust for Public Lands.
This is interesting because whereas a limited number of tract centroids failed
to overlap with the 10-minute buffer, all of the block group centroids intersect.

## Accessibility Map
Here's a map of the values. Note that there is also a shiny app in this project
that allows the user to set different values for the logsum parameters.

```{r access_map}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", c(bg_access$access_ls, bg_access$multi_ls), n = 5)
leaflet(bg_access %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "access", color = ~ pal(access_ls), stroke = 1,
              label = ~as.character(round(access_ls, 2))) %>%
  addPolygons(group = "Kinner", color = ~ pal(multi_ls), stroke = 1,
              label = ~as.character(round(multi_ls, 2))) %>%
  addLayersControl(overlayGroups = c("access", "Kinner"))
```

>>>>

---
title: "Logsums"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(sf)
library(spdep)
library(spatialreg)
library(leaflet)
library(lubridate)
library(VGAM)
```

In this document we calculate the distance between each tract in New York City 
and each park to help calculate a logsum-based measure of accessibility. We 
cleaned and gathered the data in the previous two pages.

```{r load}
open_spaces <- read_rds("data/open_spaces.rds")
tracts <- read_rds("data/tracts.rds")
```

We are going to do a point-to-polygon distance calculation. The geometric centroid
of the tracts are not satisfying, so we will use the population-weighted centroid
instead.

```{r pop_weighted_centroids}
pop_weighted_centroids <- tracts %>% as_tibble() %>%
  dplyr::select(geoid) %>%
  left_join(
    read_csv("https://www2.census.gov/geo/docs/reference/cenpop2010/tract/CenPop2010_Mean_TR36.txt") %>%
      transmute(geoid = str_c(STATEFP, COUNTYFP, TRACTCE), LATITUDE, LONGITUDE)
  ) %>%
  # convert to sf and project into the feet-based projection used in the
  # parks dataset.
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(3628) %>%
  write_rds("data/pop_weighted_centroids.rds")

```

The population-weighted centroid also serves as the orientation for the spatial
weights matrix.

```{r W}
# determine which tracts are neighbors of each other. Need to be less than
# 1.8 miles away from the centroids
tracts.dnn <- dnearneigh(as(pop_weighted_centroids, "Spatial"), 0, 1.8 * 5280)
# calculate the inverse distance to use as a weight on the neighbor connection
tract_dists <- nbdists(tracts.dnn, as(pop_weighted_centroids, "Spatial"))
write_rds(tract_dists, "data/tract_distances.rds")
```

## Distances
We want to make sure that the distance between a park and a tract centroid is
based on the edge of the polygon instead of the centroid thereof, so that spaces
at the corners of Central park, for instance, don't appear as longer than they should be.
It seems that things are working appropriately.

```{r test_measurement}
central_park <- open_spaces %>% filter(name == "Central Park")
pop_weighted_centroids$central_park <- log(as.numeric(st_distance(pop_weighted_centroids, central_park)))

dmeasure <- colorNumeric(palette = "Blues", domain = pop_weighted_centroids$central_park)
leaflet(pop_weighted_centroids %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addCircleMarkers(color = ~dmeasure(central_park))
```

Now we can calculate the distance from every park to every tract population
centroid, and write the distances and neighbors weights matrix to a file.
```{r distances}
# distances from all tract centroids to all parks
distances <- st_distance(pop_weighted_centroids, open_spaces, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()

# assert that a tract must be at least 1/10 mile from a park
distances <- pmax(distances, 0.1)
rownames(distances) <- pop_weighted_centroids$geoid
colnames(distances) <- open_spaces$id

distance_df <- distances %>%
  as_tibble(rownames = "geoid") %>%
  pivot_longer(cols = -geoid, names_to = "park_id", values_to = "distance")

write_rds(distances, "data/distances.rds")
```

## Logsums

Here's the function we use to calculate the logsums. 
```{r calculate_park_logsums}
source("R/logsums.R")
print(calculate_park_logsums)
```

We use this function with asserted beta values of -1.768 and 0.373 to calculate the
accessibility logsum.

```{r logsum_vars}
logsums <- distance_df %>%
  left_join(
    open_spaces %>% 
      transmute(
        park_id = id, acres,
        tweet_count, trails, playgrounds, courts,
      ) %>% 
      st_set_geometry(NULL),  
    by = "park_id"
  ) %>%
  mutate(
    u_access = log(distance)*-1.768 + log(acres) * 0.373,
    u_multi  = u_access + 
      trails * 0.99 +  
      playgrounds * 0.26 +  
      courts * 0.43,
    u_tweets = u_access + 0.1 * yeo.johnson(tweet_count, lambda = 0),
  ) %>%
  group_by(geoid) %>%
  summarise(
    access_ls = log(sum(exp(u_access))),
    multi_ls = log(sum(exp(u_multi))),
    tweets_ls = log(sum(exp(u_tweets)))
  ) %>%
  mutate_at(.vars = c("access_ls", "multi_ls", "tweets_ls"),  ~ (. - mean(.)) / sd(.))

logsum_vars <- tracts %>%
  left_join(logsums, by = "geoid")
```

We are also interested in looking at what happens when the tracts are within a 
10-minute walk of a park, the threshold established by the Trust for Public Lands.
```{r nyc10mw}
access_10mw <- st_read("data/nyc_10mw.geojson") %>%
  summarise(d = 1)

inside_access <- pop_weighted_centroids %>%
  st_transform(4326) %>%
  st_join(access_10mw, st_intersects) %>%
  filter(!is.na(d))

tracts_access <- tracts %>%
  left_join(logsum_vars, by = "geoid") %>%
  mutate(walk_10 = ifelse(geoid %in% inside_access$geoid, TRUE, FALSE)) %>%
  write_rds("data/tracts_access.rds")
```

Here's a map of the values. Note that there is also a shiny app in this project that
allows the user to set different values for the logsum parameters.
```{r access_map}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", c(tracts_access$access_ls, tracts_access$multi_ls), n = 5)
bin <- colorFactor("Dark2", tracts_access$walk_10)
leaflet(tracts_access) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "access", color = ~ pal(access_ls),
              label = ~as.character(round(access_ls, 2))) %>%
  addPolygons(group = "tweets", color = ~ pal(tweets_ls),
              label = ~as.character(round(tweets_ls, 2))) %>%
  addPolygons(group = "Kinner", color = ~ pal(multi_ls),
              label = ~as.character(round(multi_ls, 2))) %>%
  addPolygons(group = "10-minute walk", color = ~ bin(walk_10)) %>%
  addLayersControl(baseGroups = c("access", "tweets", "Kinner", "10-minute walk"))
```
